### 2014-04-05 doc/uist14/conclusion.tex
outlined the uist paper

In this project, we explored a novel way of interacting with physical devices by capturing user's attention. The prototype we built demonstrates the possibility of identifying user's line of sight to select and control target appliances. Furthermore, we designed a mechanism to intuitively choose from multiple targets when they are close-by.

There are still a few more components that require enhancement: (A) Thorough verifications of the IR strength, angles, and reflecting effects. (B) A more robust communication protocol with better error handling. (C) A more stable and richer input method. (D) User study.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

In this paper we have presented our system and solutions for the head orientation selection.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-05 doc/uist14/discussion.tex
outlined the uist paper

\section{Discussion}
Our study procedures demonstrated that users can successfully select and control smart appliances with head-worn infrared targeting, and that this technique outperforms list selection on the Google Glass wearable device. In this section, we revisit some of the results and observations and discuss their larger significance and potential paths for future work.

\paragraph{Meaningful Results?}
While the performance increase in our targeting study is statistically significant, readers may wonder whether it is truly meaningful. We believe it is, for two reasons: first, our technique avoids the problems of {\em naming} and {\em scoping} inherent in any interface that uses representations of objects (e.g., a list of identifiers) rather than the objects themselves. We argue that this disintermediation of interaction leads to a cognitively simpler design. Second, our existing study only showed results for a modest number of targets. Our technique should have a wider margin as the number of targets increases. Of course, the number of targets one can realistically expect may vary across application domains.

\paragraph{Hands-Free Operation}
While head movement is not as precise as hand positioning (e.g., in Patel's mobile laser pointing~\cite{patel_2-way_2003}), one key benefit of our target acquisition step is that it does not require the user's hands. This raises the question if the rest of the interaction (disambiguation and device control) could also be achieved in a hands-free fashion. Voice-command control is an obvious candidate, though such approaches have not found widespread adoption because of social acceptability and other factors.

\paragraph{Hardware Limitations}
In our prototype, the IR emitter is fixed in a single position. Due to different sizes and shapes of the users’ heads, the emitter may not line up exactly with their head orientation. An adjustable emitter (paired with a suitable calibration routing) would improve the performance of our design. Some of our users also mentioned that they had to move closer to some targets to successfully select them. A stronger emitter could overcome these problems, but care has to be taken to avoid possible reflection problems where IR light bounces off a wall and hits an unintended target behind the user.

More importantly though, the main limitation of our design is that extra hardware for infrared communication is needed for the head-mounted device and each controllable appliance. One potential approach to sidestep this requirement would be to combine the growing availability of high-resolution indoor maps with live data from the point-of-view camera on the device to determine what a user is looking at without any infrared data exchange.

\paragraph{The Midas Look}
Our participants suggested eliminating explicit initiation of a connection by the user. However, one of the design guidelines for near-eye displays is to avoid pushing information to the display without an initial request from the user --- flashing device information on screen each time a user moves their head would surely be counterproductive. Future work should investigate how to intelligently decide when and how to initiate interaction for the user.

\paragraph{Where is the Target?}
One open design question of our approach is where infrared receivers should be placed. For a light, one might put a received on the light itself, or on the light switch, to cater to existing expectations. For volume control, the infrared receiver might be located on a speaker or on the amplifier. A thorough study of user preferences would be interesting; though we also point out that our architecture could easily support multiple receivers that end up controlling the same appliance.

%Our prototype does not yet have a general way of communicating device descriptions and capabilities - this means our prototype cannot yet control new, unknown devices - they first have to be modeled. We may extend the PUC work~\cite{nichols} to address this issue.

%bjoern: these below are details about what happened in the experiment; they are less relevant to the research contributions of the paper.
%\item During the experiment, a lot of wireless messages were transferred for communication and logging purposes. Some messages got lost or delayed and caused the users had to redo that particular action. Infrared mode requires a lot of more handshake communications than the list mode. This means that in a more robust infrastructure, the performance of infrared mode can improve more than that of the list mode.

%\item In list mode, users’ eyes focuses more on the display to traverse between the list while in infrared mode, users’ eyes remain in the physical world more.

%\item The list mode could also be improved by making the swiping gestures more responsive, e.g. swiping speed affects traversing speed. % too detailed - partially covered later

%\item In our experiment, after users locates the nodes by ID, they can instantly figure out their names by the label provided. In realistic environments, users have to recall the names of the appliances, which would take a lot more time. Furthermore, finding names such as “Living room corner stand lamp” would be more difficult than an alphabet. On the other hand, infrared mode will most likely to be unaffected since it does not require name-recalling or text matching. %too detailed

>>>>

\section{Discussion}
\label{sec:discussion}

We discuss a few issues with the system.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-05 doc/uist14/evaluation.tex
outlined the uist paper

%\textcolor{red}{Will fill this evaluation when we get to it.}
Constrained by the time within a semester, we haven't conducted a formal user study. Based on our discussions with the instructors, we have some initial plans as to how to work on evaluation -- (A) Make a mobile (web) app that can remotely control the appliances through the gateway. Give a sequence of tasks to the tester and have him complete the tasks first on smartphone then use the Glass. (B) Observe the differences between using LED, Recon screen, and other feedbacks such as audio, and how that affects the user experience.

We are planning to have evaluation done within the summer.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

This section we present our work on the evaluation and user study.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-05 doc/uist14/intro.tex
outlined the uist paper

Increasingly, objects in our built environment are equipped with computation and communication, making them potential targets for user interaction. Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for advertising or context-aware computing \bjoern{second part needs to be stronger}. Both uses rely on {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.

Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}. While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed (which also limits use to explicit scenarios); that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at targets in the world.

Emerging head-worn computing devices have the potential to overcome some of these limitations: because they are worn, they do not require retrieval; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to overlay information. \bjoern{needs a bridge from head-worn device to using head orientation as a feature}.

In this paper, we investigate how to leverage head-orientation for target selection in physical spaces. We characterize this interaction modality because it can readily be implemented with small hardware changes to emerging wearable devices. We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose. We use infrared (IR) communication between Glass and target appliances to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.  
%Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio. Target appliances similarly have IR receivers and radios.

Users look in the direction of a target they wish to acquire to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume), then tap to confirm selection.  
%
Head orientation is a useful but imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus. Therefore, it requires area selection \bjoern{reference} combined with disambiguation techniques.
In practice, our IR beam's selection area has a diameter of 60-120cm at distances up to 5m.
If multiple targets fall within communication range, targets compare received IR light intensity readings to disambiguate. Users can override this default selection with an  on-screen target list.

%This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target appliances while issuing control commands.

While prior work has tended to focus on proofs-of-concept, we also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control. We first report measurements of range and beam characteristics of our controller. We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface. We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller. 
We then quantify the additional benefit of using IR-intensity disambiguation \bjoern{and find XYZ.}
%
\bjoern{Say something about the inferred map if we get around to it}.

We then demonstrate how the technique can be applied to a "universal remote control" scenario. In our implementation, user look at a target appliance and tap to load an appliance-specific control UI on the head-mounted display, which enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}) \bjoern{we may need a different figure 1 with the changed focus of the paper}. We also report qualitative results from users performing  home automation tasks with this system.

@@

>>>>

%% from the swarm vision to the necessity of selection
Increasingly, objects in our built environment are equipped with computation and communication, making them potential targets for user interaction. Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for targeted-advertisement or context-aware computing \bjoern{second part needs to be stronger}. Both uses rely on {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.

%% previous approaches are limited
Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}. While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed (which also limits use to explicit scenarios); that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at targets in the world.

%% introducing head-worn computing and head orientation
Emerging head-worn computing devices have the potential to overcome some of these limitations: because they are worn, they do not require retrieval; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to overlay information. Also such computing devices are well aligned with the user's line of sight, which is a strong indication of the user's interaction targets \ben{citation needed here}.

%% head orientation limit
Though natural, head orientation is an imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus. 
% should we argue gaze tracking doesn't help much but only increases the complexity?
Therefore, it requires area selection \bjoern{reference} combined with disambiguation techniques.

%% the inaccuracy of head orientation
In this paper, we investigate how to leverage head-orientation for target selection in physical spaces. We propose a simple low-cost solution that can be readily implemented with small hardware changes to emerging wearable devices. Specifically, we augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose. We use infrared (IR) communication between Glass and target appliances to capture the user's head orientation. Because IR signal usually covers a cone shape of area (a diameter of 60-120cm and distance up to 6m), it's suitable for the area selection modality. To achieve the disambiguation when multiple targets have received IR signals, we propose three different ways of disambiguation. First, the targets compare the IR signal intensity and we propose some intuitive heuristic to infer the user's intention in disambiguation. Second, we incorporate the sensors on Google Glass to track the user's head movement and combine this with IR intensity changes. Last but not least, we always leave the option of manually changing the current targets to the user such but limit the scope of potential targets to the ones that have received IR signals.
%% three ways

While prior work has tended to focus on proofs-of-concept for head-orientation based interaction, we also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control. We first report measurements of range and beam characteristics of our controller. We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface. We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller. 
We then quantify the additional benefit of using IR-intensity disambiguation \bjoern{and find XYZ.}. \bjoern{Say something about the inferred map if we get around to it}.

Such head orientation-based selection enables a wide range of context-aware applications. Examples include smart home remote control, break reminder monitor starer, museum attention tracking, indoor positioning, etc. In Figure\,\ref{fig:teaser}, it's a demonstration of the ``universal remote control'' scenario. The user can easily select the smart appliances by simply looking at it's general direction and confirm such selection with either voice command or by tapping the Glass input pad. Then an appliance-specific control UI will be shown on the head-mounted display. For this application, we have asked 14 participants to try the system and we report the qualitative results from them performing home automation tasks.

@@

%% Contribution
In this paper, by presenting our work on head orientation-based beam cursor,  we make the following contributions:
\begin{itemize}
\item We design and implement a low-cost, easy-to-use head orientation system that enables interaction with physical environment.
\item To solve the inherent inaccuracy of head orientation, we have proposed three different ways to achieve disambiguation and we evaluate them based on real-world user study.
\item We also demonstrate that several useful applications can be easily built with the head orientation system.
\end{itemize}

\ben{Having problem fixing the reference}.
In the remainder of this paper, we will first describe the related works in physical selection and targeting. Given that we focus our scope on head orientation, in Section~\ref{sec:background} we will briefly review human's neck ergonomics as the background for head orientation. In Section~\ref{sec:syst-design-prot}, we then present our system designed for the study of head orientation based selection. The prototype is also what we have used for building the example applications. The disambiguation techniques are discussed in Section~\ref{sec:disamb-techn}, which is followed by the study and evaluation in Section~\ref{sec:evaluation}. To further show the usefulness of having such head orientation-based selection, we describe the enabled applications in Section~\ref{sec:applications}, with detailed implementation about the ``universal remote control'' application. The discussion and conclusion are in Section~\ref{sec:discussion} and Section~\ref{sec:conclusion} respectively.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:
